---
title: "p8105_nyc_shooting"
output: html_document
date: "2024-11-19"
---
```{r, message=FALSE}
library(sf)
library(dplyr)
library(readr)
library(stringr)
library(lubridate)
library(timeDate)
library(suncalc)
library(purrr)
library(readxl)
library(tidyverse)

```
# This part focus on data cleaning, tidying and merging.
## adding neighborhood information to the shooting dataset
```{r, message=FALSE}
df=read_csv("NYPD_Shooting_Incident_Data__Historic__20241119.csv")

# Ensure coordinate columns are numeric
df = df %>%
  mutate(
    X_COORD_CD = as.numeric(X_COORD_CD),
    Y_COORD_CD = as.numeric(Y_COORD_CD)
  )

# Convert to sf object with CRS EPSG:2263
df_sf = st_as_sf(df, coords = c("X_COORD_CD", "Y_COORD_CD"), crs = 2263)

# Download and read the GeoJSON file
download.file(
  url = "https://data.insideairbnb.com/united-states/ny/new-york-city/2024-09-04/visualisations/neighbourhoods.geojson",
  destfile = "neighbourhoods.geojson",
  mode = "wb"
)
neighbourhoods = st_read("neighbourhoods.geojson")

# Check and transform CRS of neighborhoods
if (st_crs(neighbourhoods)$epsg != 2263) {
  neighbourhoods = st_transform(neighbourhoods, crs = 2263)
}

# Perform the spatial join
df_joined = st_join(df_sf, neighbourhoods, left = TRUE)

# Add neighborhood information to the data frame
df$Neighborhood = df_joined$neighbourhood

# Download and use the neighborhoods CSV file
download.file(
  url = "https://data.insideairbnb.com/united-states/ny/new-york-city/2024-09-04/visualisations/neighbourhoods.csv",
  destfile = "neighbourhoods.csv",
  mode = "wb"
)
neighborhoods_csv = read_csv("neighbourhoods.csv")

# Merge additional attributes if necessary
df = df %>%
  left_join(neighborhoods_csv, by = c("Neighborhood" = "neighbourhood"))

```
## add nta data
```{r}
df$X_COORD_CD <- as.numeric(df$X_COORD_CD)
df$Y_COORD_CD <- as.numeric(df$Y_COORD_CD)

# Convert to sf object
df_sf <- st_as_sf(df, coords = c("X_COORD_CD", "Y_COORD_CD"), crs = 2263)

# Read neighborhood shapefile
neighborhoods <- st_read("./nynta2020_24d/nynta2020.shp")

# Transform CRS if necessary
neighborhoods <- st_transform(neighborhoods, crs = st_crs(df_sf))

# Perform spatial join
df_joined <- st_join(df_sf, neighborhoods, left = TRUE)

# Add neighborhood information to your original data frame
df$NTA <- df_joined$NTAName  # Adjust based on actual column name
```

## add 2010 nta data
```{r}
df$X_COORD_CD <- as.numeric(df$X_COORD_CD)
df$Y_COORD_CD <- as.numeric(df$Y_COORD_CD)

# Convert to sf object
df_sf <- st_as_sf(df, coords = c("X_COORD_CD", "Y_COORD_CD"), crs = 2263)

# Read neighborhood shapefile
neighborhoods_2010 <- st_read("./nynta2010_24d/nynta2010.shp")

# Transform CRS if necessary
neighborhoods_2010 <- st_transform(neighborhoods_2010, crs = st_crs(df_sf))

# Perform spatial join
df_joined <- st_join(df_sf, neighborhoods_2010, left = TRUE)

# Add neighborhood information to your original data frame
df$NTA_2010 <- df_joined$NTAName  # Adjust based on actual column name
```

## add time information
for future study, i need to add variables related to date and time of the gunshot incidence.

### decide whether the date of shooting is national holiday
```{r}
df$OCCUR_DATE <- as.Date(df$OCCUR_DATE, format = "%m/%d/%Y")

# Define the range of years, handling NA values
years <- seq(
  year(min(df$OCCUR_DATE, na.rm = TRUE)),
  year(max(df$OCCUR_DATE, na.rm = TRUE)),
  by = 1
)
us_holidays = holidayNYSE(years)
us_holidays = as.Date(us_holidays)
# Add a new column indicating whether the date is a holiday, year and month information of a date
df = df %>%
  mutate(
    Is_Holiday = OCCUR_DATE %in% us_holidays,
    Year = year(OCCUR_DATE),
    Month = month(OCCUR_DATE)
  )

```
### decide whether the sky is dark at the shooting time, using the dawn and dusk time:
```{r}
df$OCCUR_DATE <- as.character(df$OCCUR_DATE)
df$OCCUR_TIME <- as.character(df$OCCUR_TIME)

# Combine OCCUR_DATE and OCCUR_TIME into a single datetime string
datetime_str <- paste(df$OCCUR_DATE, df$OCCUR_TIME)

# Parse datetime using lubridate
df$OCCUR_DATETIME <- ymd_hms(datetime_str, tz = "America/New_York")
# Extract date from OCCUR_DATETIME
df$DATE <- as.Date(df$OCCUR_DATETIME)

# Get unique dates
unique_dates <- unique(df$DATE)
latitude <- 40.7128
longitude <- -74.0060
# Calculate dawn and dusk times
sun_times <- getSunlightTimes(
  date = unique_dates,
  lat = latitude,
  lon = longitude,
  keep = c("dawn", "dusk"),
  tz = "America/New_York"
)

# Merge with the original dataframe
df <- df %>%
  left_join(sun_times[, c("date", "dawn", "dusk")], by = c("DATE" = "date"))

# Determine if the sky is dark considering twilight
df <- df %>%
  mutate(
    Sky_Is_Dark = OCCUR_DATETIME < dawn | OCCUR_DATETIME >= dusk
  )
df <- df %>%
  select(-DATE, -dawn, -dusk)
```

## adding total population data of each NTA

```{r}
popu <- suppressWarnings(
  read_excel("nyc_detailed-race-and-ethnicity-data_2020_core-geographies.xlsx", 
             sheet = 1,
             range = "A4:I2599",
             col_types = c("numeric", "text", "text", "text", "numeric", "text", "text", "numeric", "numeric")) |>
    filter(`Orig Order` >= 2334 & `Orig Order` <= 2595) |>
    select(GeoName, NTAType, Pop) |>
    rename(Total_population = Pop)
)

df <- df |>
  left_join(popu, by = c("NTA" = "GeoName")) |>
  mutate(
    NTAType = case_when(
      NTAType == 0 ~ "Residential",
      NTAType == 9 ~ "Park",
      NTAType == 8 ~ "Airport",
      NTAType == 7 ~ "Cemetery",
      NTAType == 6 ~ "Other Special Areas",
      NTAType == 5 ~ "Rikers Island",
      is.na(NTAType) ~ "Unknown")
  )

```

## adding cdta data

```{r}
df$X_COORD_CD <- as.numeric(df$X_COORD_CD)
df$Y_COORD_CD <- as.numeric(df$Y_COORD_CD)

df_sf <- st_as_sf(df, coords = c("X_COORD_CD", "Y_COORD_CD"), crs = 2263)

cdta <- st_read("./nycdta2020_24d/nycdta2020.shp")

cdta <- st_transform(cdta, crs = st_crs(df_sf))

df_joined <- st_join(df_sf, cdta, left = TRUE)

df$CDTA <- df_joined$CDTA2020

df$CDTA <- gsub("([A-Z]{2})(\\d{2})", "\\1 \\2", df$CDTA)

```

## add poverty
```{r}
neighborhood_poverty <- read.csv("./data/neighborhood_poverty.csv")

# Select specific columns from neighborhood_poverty
neighborhood_poverty_selected <- neighborhood_poverty %>% 
  filter(TimePeriod == '2017-21') %>% 
  filter(GeoType == 'NTA2020' ) %>% 
  select(Number, Percent, Geography)
# Merge the dataframes on 'NTA' and 'Geography', keeping all information from data_processing
df_poverty <- left_join(df, neighborhood_poverty_selected, by = c("NTA" = "Geography"))

# Rename columns after merging
df_poverty <- df_poverty %>% 
  rename(Number_poverty = Number, Percent_poverty = Percent)
```

## add education
```{r}
neighborhood_education=read.csv("./data/graduated_high_school.csv")
neighborhood_education_selected <- neighborhood_education %>% 
  filter(TimePeriod == '2017-21') %>% 
  filter(GeoType == 'NTA2020' ) %>% 
  select(Number, Percent, Geography)
# Merge the dataframes on 'NTA' and 'Geography', keeping all information from data_processing
df_education <- left_join(df_poverty, neighborhood_education_selected, by = c("NTA" = "Geography"))

# Rename columns after merging
df_education <- df_education %>% 
  rename(Number_education = Number, Percent_education = Percent)
# filter between 2017 and 2023
df_education$OCCUR_DATE <- as.Date(df_education$OCCUR_DATE, format = "%Y-%m-%d")

# Filter the incidents that happened between 2017 and 2023
df_education <- df_education %>% 
  filter(OCCUR_DATE >= as.Date("2017-01-01") & OCCUR_DATE <= as.Date("2023-12-31"))
```

```{r}
#add a column that shows the shooting incident rate of NTAs in that year
df_education <- df_education %>%
  group_by(NTA, Year) %>%
  mutate(incident_rate_by_year_nta = (n() / Total_population)*100) %>%
  ungroup()

#add a column that shows the shooting incident rate of boroughs in that year
boro_population <- df_education %>%
  group_by(BORO, Year) %>%
  summarise(total_population_boro = sum(Total_population, na.rm = TRUE)) %>%
  ungroup()

df_education <- df_education %>%
  left_join(boro_population, by = c("BORO", "Year")) %>%
  group_by(BORO, Year) %>%
  mutate(incident_rate_by_year_boro = (n() / total_population_boro)*100) %>%
  ungroup()

#Rename a column
df_education <- df_education %>%
  rename(Total_population_nta = Total_population)

write.csv(df_education, "data_final.csv", row.names = FALSE)
```