---
title: "descriptive_statistics"
output: html_document
date: "2024-12-03"
---

# Descriptive Statistics of NYC Shooting Data
```{r}
library(ggplot2)
library(readxl)
library(readr)
library(dplyr)
library(knitr)
library(tidyr)
```


```{r}
# Load the dataset
df_descriptive=read_csv("data_final.csv")
```
## time

### Plotting the line chart for shooting incidents over time

```{r}
# Converting OCCUR_DATE to Date type
df_descriptive$OCCUR_DATE <- as.Date(df_descriptive$OCCUR_DATE, format = "%Y-%m-%d")

# Grouping the data by date to count the number of incidents
date_counts <- as.data.frame(table(df_descriptive$OCCUR_DATE))
colnames(date_counts) <- c('Date', 'Count')
date_counts$Date <- as.Date(date_counts$Date)

# Plotting the line chart for shooting incidents over time
ggplot(date_counts, aes(x = Date, y = Count)) +
  geom_line(color = 'blue') +
  labs(title = 'Shooting Incidents Over Time in NYC',
       x = 'Date',
       y = 'Number of Shooting Incidents') +
  theme_minimal()
```

### Plotting the line chart for shooting incidents by year with data points labeled

```{r}
# Grouping the data by year to count the number of incidents
df_descriptive$Year <- as.factor(df_descriptive$Year)
year_counts <- as.data.frame(table(df_descriptive$Year))
colnames(year_counts) <- c('Year', 'Count')


# Plotting the line chart for shooting incidents by year with data points labeled
ggplot(year_counts, aes(x = Year, y = Count, group = 1)) +
  geom_line(color = 'blue') +
  geom_point(size = 3) +
  geom_text(aes(label = Count), vjust = -0.5, size = 3) +
  labs(title = 'Shooting Incidents by Year in NYC',
       x = 'Year',
       y = 'Number of Shooting Incidents') +
  theme_minimal()
```

### Plotting the bar chart for shooting incidents by month

```{r}
# Grouping the data by month
df_descriptive$Month <- as.factor(df_descriptive$Month)
month_counts <- as.data.frame(table(df_descriptive$Month))
colnames(month_counts) <- c('Month', 'Count')

# Plotting the bar chart for shooting incidents by month
ggplot(month_counts, aes(x = Month, y = Count)) +
  geom_bar(stat = 'identity', fill = 'skyblue') +
  geom_text(aes(label = Count), vjust = -0.5) +
  labs(title = 'Bar Chart of Shooting Incidents by Month in NYC',
       x = 'Month',
       y = 'Number of Shooting Incidents') +
  theme_minimal()
```

```{r}
# Convert OCCUR_TIME to a proper time format
df_descriptive$OCCUR_TIME <- format(strptime(df_descriptive$OCCUR_TIME, format = "%H:%M:%S"), "%H:%M:%S")

# Plotting the density of the occurrence time
ggplot(df_descriptive, aes(x = as.POSIXct(OCCUR_TIME, format = "%H:%M:%S"))) +
  geom_density(fill = 'skyblue', alpha = 0.6) +
  labs(title = 'Density Plot of Shooting Incident Occurrence Time in NYC',
       x = 'Time of Occurrence',
       y = 'Density') +
  scale_x_datetime(date_labels = "%H:%M") +
  theme_minimal()
```

## geo

### Table showing Top 10 total incident NTA in Each Borough

```{r}

# Table showing Top 10 total incident NTA in Each Borough
incident_by_nta_borough <- df_descriptive %>%
  drop_na() %>%
  group_by(BORO, NTA) %>%
  summarise(total_incidents = n()) %>%
  arrange(BORO, desc(total_incidents)) %>%
  group_by(BORO) %>%
  slice_max(n = 10, order_by = total_incidents) %>%
  select(-total_incidents) %>%
  group_by(BORO) %>%
  mutate(row_num = row_number()) %>%
  pivot_wider(names_from = BORO, values_from = NTA) %>%
  unnest(cols = c(BRONX, BROOKLYN, MANHATTAN, QUEENS, `STATEN ISLAND`)) %>%
  select(row_num, everything())%>%
  slice(1:10)

# Display the table in the desired format
kable(incident_by_nta_borough, caption = "Top 10 NTAs with Total Shooting Incidents in Each Borough (2017-2023)")
```

### Table showing Top 10 NTAs with the Highest Total Shooting Incidents

```{r}
# Table showing Top 10 NTAs with the Highest Total Shooting Incidents
top_10_nta_overall <- df_descriptive %>%
  group_by(BORO, NTA) %>%
  summarise(total_incidents = n(),
            total_population = first(Total_population_nta),
            percent_poverty = first(Percent_poverty),
            percent_education = first(Percent_education)) %>%
  arrange(desc(total_incidents)) %>%
  ungroup() %>%
  slice(1:10) %>%
  select(BORO, NTA, total_incidents, total_population, percent_poverty, percent_education)
# Display the table in the desired format
kable(top_10_nta_overall, digits = 2, caption = "Top 10 NTAs with the Highest Total Shooting Incidents")
```

#### MANHATTAN

```{r}

# Table showing Top 10 NTAs with the Highest Total Shooting Incidents in MANHATTAN
top_10_nta_overall <- df_descriptive %>%
  filter(BORO == 'MANHATTAN') %>% 
  group_by(BORO, NTA) %>%
  summarise(total_incidents = n(),
            total_population = first(Total_population_nta),
            percent_poverty = first(Percent_poverty),
            percent_education = first(Percent_education)) %>%
  arrange(desc(total_incidents)) %>%
  ungroup() %>%
  slice(1:10) %>%
  select(BORO, NTA, total_incidents, total_population, percent_poverty, percent_education)
# Display the table in the desired format
kable(top_10_nta_overall, digits = 2, caption = "Top 10 NTAs with the Highest Total Shooting Incidents")
```

#### MANHATTAN

```{r}

# Table showing Top 10 NTAs with the Highest Total Shooting Incidents in MANHATTAN
top_10_nta_overall <- df_descriptive %>%
  filter(BORO == 'MANHATTAN') %>% 
  group_by(BORO, NTA) %>%
  summarise(total_incidents = n(),
            total_population = first(Total_population_nta),
            percent_poverty = first(Percent_poverty),
            percent_education = first(Percent_education)) %>%
  arrange(desc(total_incidents)) %>%
  ungroup() %>%
  slice(1:10) %>%
  select(BORO, NTA, total_incidents, total_population, percent_poverty, percent_education)
# Display the table in the desired format
kable(top_10_nta_overall, digits = 2, caption = "Top 10 NTAs with the Highest Total Shooting Incidents in MANHATTAN")
```

#### BROOKLYN

```{r}

# Table showing Top 10 NTAs with the Highest Total Shooting Incidents in BROOKLYN
top_10_nta_overall <- df_descriptive %>%
  filter(BORO == 'BROOKLYN') %>% 
  group_by(BORO, NTA) %>%
  summarise(total_incidents = n(),
            total_population = first(Total_population_nta),
            percent_poverty = first(Percent_poverty),
            percent_education = first(Percent_education)) %>%
  arrange(desc(total_incidents)) %>%
  ungroup() %>%
  slice(1:10) %>%
  select(BORO, NTA, total_incidents, total_population, percent_poverty, percent_education)
# Display the table in the desired format
kable(top_10_nta_overall, digits = 2, caption = "Top 10 NTAs with the Highest Total Shooting Incidents in BROOKLYN")
```

#### BRONX

```{r}

# Table showing Top 10 NTAs with the Highest Total Shooting Incidents in BROOKLYN
top_10_nta_overall <- df_descriptive %>%
  filter(BORO == 'BRONX') %>% 
  group_by(BORO, NTA) %>%
  summarise(total_incidents = n(),
            total_population = first(Total_population_nta),
            percent_poverty = first(Percent_poverty),
            percent_education = first(Percent_education)) %>%
  arrange(desc(total_incidents)) %>%
  ungroup() %>%
  slice(1:10) %>%
  select(BORO, NTA, total_incidents, total_population, percent_poverty, percent_education)
# Display the table in the desired format
kable(top_10_nta_overall, digits = 2, caption = "Top 10 NTAs with the Highest Total Shooting Incidents in BRONX")
```

#### QUEENS

```{r}

# Table showing Top 10 NTAs with the Highest Total Shooting Incidents in BROOKLYN
top_10_nta_overall <- df_descriptive %>%
  filter(BORO == 'QUEENS') %>% 
  group_by(BORO, NTA) %>%
  summarise(total_incidents = n(),
            total_population = first(Total_population_nta),
            percent_poverty = first(Percent_poverty),
            percent_education = first(Percent_education)) %>%
  arrange(desc(total_incidents)) %>%
  ungroup() %>%
  slice(1:10) %>%
  select(BORO, NTA, total_incidents, total_population, percent_poverty, percent_education)
# Display the table in the desired format
kable(top_10_nta_overall, digits = 2, caption = "Top 10 NTAs with the Highest Total Shooting Incidents in QUEENS")
```

#### STATEN ISLAND

```{r}

# Table showing Top 10 NTAs with the Highest Total Shooting Incidents in BROOKLYN
top_10_nta_overall <- df_descriptive %>%
  filter(BORO == 'STATEN ISLAND') %>% 
  group_by(BORO, NTA) %>%
  summarise(total_incidents = n(),
            total_population = first(Total_population_nta),
            percent_poverty = first(Percent_poverty),
            percent_education = first(Percent_education)) %>%
  arrange(desc(total_incidents)) %>%
  ungroup() %>%
  slice(1:10) %>%
  select(BORO, NTA, total_incidents, total_population, percent_poverty, percent_education)
# Display the table in the desired format
kable(top_10_nta_overall, digits = 2, caption = "Top 10 NTAs with the Highest Total Shooting Incidents in STATEN ISLAND")
```

### Table showing Top 10 NTAs with the Highest Incident Rate

```{r}
# Get the Top 10 NTAs with the Highest Incident Rate by Year
top_10_incident_rate <- df_descriptive %>%
  distinct(incident_rate_by_year_nta, .keep_all = TRUE) %>%
  filter(NTAType == 'Residential') %>%
  arrange(desc(incident_rate_by_year_nta)) %>%
  slice(1:10)

# Pull information for the Top 10 NTAs
top_10_nta_incident_rate <- top_10_incident_rate %>%
  select(NTA, BORO, incident_rate = incident_rate_by_year_nta,
         total_population = Total_population_nta,
         percent_poverty = Percent_poverty,
         percent_education = Percent_education)

# Display the table in the desired format
kable(top_10_nta_incident_rate, digits = 6, caption = "Top 10 NTAs with the Highest Incident Rate by Year in Each Borough, Including Population, Poverty, and Education Data")
```

## victim
### Plot: Description of Victims

```{r}
# Plot: Description of Victims
victim_summary <- df_descriptive %>%
  filter(VIC_SEX %in% c("M", "F")) %>%  # Filter out unknown
  group_by(VIC_SEX, VIC_AGE_GROUP, VIC_RACE) %>%
  summarise(total_victims = n()) %>%
  drop_na()


# Plotting the description of male victims
ggplot(victim_summary %>% filter(VIC_SEX == "M"), aes(x = VIC_AGE_GROUP, y = total_victims, fill = VIC_RACE)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = "Description of Male Shooting Victims by Age Group and Race",
       x = "Victim Age Group",
       y = "Number of Victims",
       fill = "Victim Race") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plotting the description of female victims
ggplot(victim_summary %>% filter(VIC_SEX == "F"), aes(x = VIC_AGE_GROUP, y = total_victims, fill = VIC_RACE)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = "Description of Female Shooting Victims by Age Group and Race",
       x = "Victim Age Group",
       y = "Number of Victims",
       fill = "Victim Race") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Plot: Bar chart showing all victim's age

```{r}
# Plot: Bar chart showing victim's age
ggplot(df_descriptive %>% filter(!is.na(VIC_AGE_GROUP)& VIC_AGE_GROUP != 1022), aes(x = VIC_AGE_GROUP)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Bar Chart of Victim's Age",
       x = "Victim Age Group",
       y = "Count of Victims") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 