---
title: "descriptive_statistics"
output: html_document
date: "2024-12-03"
---

# Descriptive Statistics of NYC Shooting Data
```{r}
library(ggplot2)
library(readxl)
library(readr)

```


```{r}
# Load the dataset
df_descriptive=read_csv("filtered_merged_dataset_sample.csv")
```

## Plotting the line chart for shooting incidents over time

```{r}
# Converting OCCUR_DATE to Date type
df_descriptive$OCCUR_DATE <- as.Date(df_descriptive$OCCUR_DATE, format = "%Y-%m-%d")

# Grouping the data by date to count the number of incidents
date_counts <- as.data.frame(table(df_descriptive$OCCUR_DATE))
colnames(date_counts) <- c('Date', 'Count')
date_counts$Date <- as.Date(date_counts$Date)

# Plotting the line chart for shooting incidents over time
ggplot(date_counts, aes(x = Date, y = Count)) +
  geom_line(color = 'blue') +
  labs(title = 'Shooting Incidents Over Time in NYC',
       x = 'Date',
       y = 'Number of Shooting Incidents') +
  theme_minimal()
```

## Plotting the line chart for shooting incidents by year with data points labeled

```{r}
# Grouping the data by year to count the number of incidents
df_descriptive$Year <- as.factor(df_descriptive$Year)
year_counts <- as.data.frame(table(df_descriptive$Year))
colnames(year_counts) <- c('Year', 'Count')


# Plotting the line chart for shooting incidents by year with data points labeled
ggplot(year_counts, aes(x = Year, y = Count, group = 1)) +
  geom_line(color = 'blue') +
  geom_point(size = 3) +
  geom_text(aes(label = Count), vjust = -0.5, size = 3) +
  labs(title = 'Shooting Incidents by Year in NYC',
       x = 'Year',
       y = 'Number of Shooting Incidents') +
  theme_minimal()
```

## Plotting the bar chart for shooting incidents by month

```{r}
# Grouping the data by month
df_descriptive$Month <- as.factor(df_descriptive$Month)
month_counts <- as.data.frame(table(df_descriptive$Month))
colnames(month_counts) <- c('Month', 'Count')

# Plotting the bar chart for shooting incidents by month
ggplot(month_counts, aes(x = Month, y = Count)) +
  geom_bar(stat = 'identity', fill = 'skyblue') +
  geom_text(aes(label = Count), vjust = -0.5) +
  labs(title = 'Bar Chart of Shooting Incidents by Month in NYC',
       x = 'Month',
       y = 'Number of Shooting Incidents') +
  theme_minimal()
```

```{r}
# Convert OCCUR_TIME to a proper time format
df_descriptive$OCCUR_TIME <- format(strptime(df_descriptive$OCCUR_TIME, format = "%H:%M:%S"), "%H:%M:%S")

# Plotting the density of the occurrence time
ggplot(df_descriptive, aes(x = as.POSIXct(OCCUR_TIME, format = "%H:%M:%S"))) +
  geom_density(fill = 'skyblue', alpha = 0.6) +
  labs(title = 'Density Plot of Shooting Incident Occurrence Time in NYC',
       x = 'Time of Occurrence',
       y = 'Density') +
  scale_x_datetime(date_labels = "%H:%M") +
  theme_minimal()