---
title: "Poverty VS Incidence Rate"
author: "Chenhui Yan"
date: "2024-12-02"
output: github_document
---
```{r}
library(dplyr)
library(plotly)
library(ggplot2)

```

## Poverty
Is there an association between the percentage of people whose income is below the poverty line and the incidence rate 1) across neighborhoods in NYC and 2) across neighborhoods in each borough?


```{r}

data_clean <- read.csv("data_final.csv")
# List all column names
names(data_clean)

# Check the structure of the data frame
str(data_clean)


# Step 1: Check the structure of your dataset to ensure the necessary columns exist
head(data_clean)

```
## Poverty across neighborhoods in NYC

```{r}
# Calculate the correlation between the poverty percentage and the incident rate
correlation <- cor(data_clean$incident_rate_by_year_nta, data_clean$Percent_poverty, use = "complete.obs")
print(paste("Correlation coefficient: ", correlation))

```
When examining all of the neighborhoods in NYC, there is a moderate positive linear relationship between the percentage of people below the poverty line (Percent_poverty) and the incident rate by neighborhood (incident_rate_by_year_nta).(r = 0.508).


```{r}
# Create a scatter plot to visualize the relationship
data_clean %>%
  plot_ly(x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA, colors = "viridis", 
          type = "scatter", mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", BORO, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) %>%
  layout(title = "Percent Below the Poverty Line and Incident Rate in NYC",
         xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
         yaxis = list(title = 'Incident Rate'),
         legend = list(title = list(text = 'Neighborhood')))
```

this plot suggests that neighborhoods with higher poverty rates tend to have higher incident rates, with some variability. This relationship may point towards socio-economic factors that influence the safety and well-being of neighborhoods in NYC.

## Poverty By Borough
```{r}
# Assuming your main data frame is named 'data_clean'

# Filter data for Manhattan
manhattan_data <- data_clean %>%
  filter(neighbourhood_group == "Manhattan")

# Filter data for Brooklyn
brooklyn_data <- data_clean %>%
  filter(neighbourhood_group == "Brooklyn")

# Filter data for The Bronx
bronx_data <- data_clean %>%
  filter(neighbourhood_group == "Bronx")

# Filter data for Staten Island
staten_island_data <- data_clean %>%
  filter(neighbourhood_group == "Staten Island")

# Filter data for Queens
queens_data <- data_clean %>%
  filter(neighbourhood_group == "Queens")

```

```{r}
# Function to compute and print correlation
compute_correlation <- function(data, borough_name) {
  correlation <- cor(
    data$incident_rate_by_year_nta,
    data$Percent_poverty,
    use = "complete.obs"
  )
  cat("Correlation coefficient for", borough_name, ":", correlation, "\n")
}

compute_correlation(manhattan_data, "Manhattan")
compute_correlation(brooklyn_data, "Brooklyn")
compute_correlation(bronx_data, "Bronx")
compute_correlation(staten_island_data, "Staten Island")
compute_correlation(queens_data, "Queens")

```

### Manhattan
```{r}
compute_correlation(manhattan_data, "Manhattan")
# Function to create scatter plot with trend line
# Scatter plot for Manhattan
data_clean |> 
  filter(neighbourhood_group == "Manhattan") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "viridis", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Manhattan",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```
### Brooklyn
```{r}
# Scatter plot for Brooklyn
data_clean |> 
  filter(neighbourhood_group == "Brooklyn") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "plasma", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Brooklyn",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```
### The Bronx
```{r}
# Scatter plot for The Bronx
data_clean |> 
  filter(neighbourhood_group == "Bronx") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "magma", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in The Bronx",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```
### Queens
```{r}
# Scatter plot for Queens
data_clean |> 
  filter(neighbourhood_group == "Queens") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "inferno", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Queens",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```
### Staten Island
```{r}
# Scatter plot for Staten Island
data_clean |> 
  filter(neighbourhood_group == "Staten Island") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "inferno", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Staten Island",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```


