---
title: "Poverty VS Incidence Rate"
author: "Chenhui Yan"
date: "2024-12-02"
output: html_document
---
```{r}
library(dplyr)
library(plotly)
library(ggplot2)

```

## Poverty
Is there an association between the percentage of people whose income is below the poverty line and the incidence rate 1) across neighborhoods in NYC and 2) across neighborhoods in each borough?


```{r}

data_clean <- read.csv("data_final.csv")
# List all column names
names(data_clean)

# Check the structure of the data frame
str(data_clean)


# Step 1: Check the structure of your dataset to ensure the necessary columns exist
head(data_clean)

```
## Poverty across neighborhoods in NYC

```{r}
# Calculate the correlation between the poverty percentage and the incident rate
correlation <- cor(data_clean$incident_rate_by_year_nta, data_clean$Percent_poverty, use = "complete.obs")
print(paste("Correlation coefficient: ", correlation))

```
When examining all of the neighborhoods in NYC, there is a moderate positive linear relationship between the percentage of people below the poverty line (Percent_poverty) and the incident rate by neighborhood (incident_rate_by_year_nta).(r = 0.508).


```{r}
# Create a scatter plot to visualize the relationship
data_clean %>%
  plot_ly(x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA, colors = "viridis", 
          type = "scatter", mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", BORO, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) %>%
  layout(title = "Percent Below the Poverty Line and Incident Rate in NYC",
         xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
         yaxis = list(title = 'Incident Rate'),
         legend = list(title = list(text = 'Neighborhood')))
```

this plot suggests that neighborhoods with higher poverty rates tend to have higher incident rates.

## Poverty By Borough
```{r}
# Assuming your main data frame is named 'data_clean'

# Filter data for Manhattan
manhattan_data <- data_clean %>%
  filter(neighbourhood_group == "Manhattan")

# Filter data for Brooklyn
brooklyn_data <- data_clean %>%
  filter(neighbourhood_group == "Brooklyn")

# Filter data for The Bronx
bronx_data <- data_clean %>%
  filter(neighbourhood_group == "Bronx")

# Filter data for Staten Island
staten_island_data <- data_clean %>%
  filter(neighbourhood_group == "Staten Island")

# Filter data for Queens
queens_data <- data_clean %>%
  filter(neighbourhood_group == "Queens")

```

```{r}
# Function to compute and print correlation
compute_correlation <- function(data, borough_name) {
  correlation <- cor(
    data$incident_rate_by_year_nta,
    data$Percent_poverty,
    use = "complete.obs"
  )
  cat("Correlation coefficient for", borough_name, ":", correlation, "\n")
}

compute_correlation(manhattan_data, "Manhattan")
compute_correlation(brooklyn_data, "Brooklyn")
compute_correlation(bronx_data, "Bronx")
compute_correlation(staten_island_data, "Staten Island")
compute_correlation(queens_data, "Queens")

```

* Positive Correlation Across All Boroughs:
 * The correlation coefficients for all boroughs are positive, indicating that as the percentage of people below the poverty line increases, the incident rate also tends to increase.
* Differences in Correlation Strength:
 * Brooklyn (0.5686) and Staten Island (0.5576) show the highest correlations, suggesting that the relationship between poverty and incident rate is stronger in these boroughs.
 * Manhattan (0.3345) and Queens (0.2797) show weaker correlations, indicating that the relationship between poverty and incident rate is less pronounced in these areas.
 * Bronx (0.4463) has a moderate correlation, which also points to a positive but not as strong relationship.

### Manhattan
```{r}
compute_correlation(manhattan_data, "Manhattan")
# Function to create scatter plot with trend line
# Scatter plot for Manhattan
data_clean |> 
  filter(neighbourhood_group == "Manhattan") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "viridis", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Manhattan",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```

Some neighborhoods, like East Harlem (North), East Harlem (South), and Chinatown-Two Bridges, seem to have both higher poverty percentages and higher incident rates, suggesting a more direct connection between poverty and incidents.

### Brooklyn
```{r}
# Scatter plot for Brooklyn
data_clean |> 
  filter(neighbourhood_group == "Brooklyn") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "plasma", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Brooklyn",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```

Brownsville, which has a high poverty rate (between 30-50%), also has some of the highest incident rates in Brooklyn, indicating that poverty may significantly impact incident rates in this neighborhood.

### The Bronx
```{r}
# Scatter plot for The Bronx
data_clean |> 
  filter(neighbourhood_group == "Bronx") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "magma", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in The Bronx",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```

Claremont Village-Claremont (East) and Fordham Heights seem to have higher incident rates compared to other neighborhoods with similar poverty levels

### Queens
```{r}
# Scatter plot for Queens
data_clean |> 
  filter(neighbourhood_group == "Queens") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "inferno", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Queens",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```

Corona and East Elmhurst show relatively high incident rates, particularly compared to neighborhoods with similar poverty percentages. 

### Staten Island
```{r}
# Scatter plot for Staten Island
data_clean |> 
  filter(neighbourhood_group == "Staten Island") |> 
  plot_ly(data = _, x = ~Percent_poverty, y = ~incident_rate_by_year_nta, 
          color = ~NTA,
          colors = "inferno", 
          type = "scatter",
          mode = "markers",
          text = ~paste("Neighborhood: ", NTA, "<br>Borough: ", neighbourhood_group, 
                        "<br>% Below Poverty Line: ", Percent_poverty, 
                        "<br>Incident Rate: ", incident_rate_by_year_nta)) |> 
    layout(title = "Percent Below the Poverty Line and Incident Rate in Staten Island",
           xaxis = list(title = 'Percentage of People Whose Income is Below the Poverty Line'),
           yaxis = list(title = 'Incident Rate'),
           legend = list(title = list(text = 'Neighborhood')))

```

In Staten Island, the relationship between poverty and incident rates is less pronounced compared to other boroughs. 

